---
title: "Facet plots"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
knitr::opts_knit$set(root.dir = 'C:/GitHub/social_care_open_data')
```

#Introduction

Going to have a wee play trying to plot facetted (small multiple) data with main dataset in the background. Based on [this plot](link here!!!!)


##Load data

I'll use the 2012 social care survey data cleaned in `import_and_tidy.Rmd`

```{r load_data}
load("produced_data/created_objects/soc_care12.rds")
```

##Load packages

```{r packages, message=FALSE}
library(tidyverse)
library(forcats)
library(ggthemes)
library(geofacet)
library(extrafont)
theme_set(theme_tufte(base_size = 18))
ubdc_palette <- c("#13AFD6", "#E6E600", "#F07329", "#35B14E", "#D7509A", "#2165AF",
                  "#BCD032","#866BAC", "#545A5D",  "#7A8082", "#E2D988", "#628DB7",
                  "#929B9A", "#93B8DA", "#31649B", "#FBF8D0", "#ACB2B4", "#D1DAE2")
```


#Plot `age_grp`

##Glimpse

Quick reminder of what the data looks like

```{r glimpse}
glimpse(soc_care12)
```


##Create data frame without grouping variable

To start with I am going to try and plot the distribution of `age_grp` using `gender` as a facet.

To do this I will need a dataframe *without* the `gender` variable

```{r ungrouped_set}
nogen <- soc_care12 %>% select (-gender)
```

...and now plot

##plot 

```{r plot_background}
ggplot(soc_care12, aes(age_grp)) + 
  geom_bar(data = nogen, fill = ubdc_palette[18]) +
  geom_bar() +
  facet_wrap(~ gender)
rm(nogen)
```

Will try something a bit more adventurous now. 

#Plot `home_care_hrs`

In `home_care_plot.md` I compared the number of home care hours across Local Authorities. Will try and see if the background plot makes this comparison easier. 

##Create dataframe

I'll need to recreate the summary data

1st of all collapse the total_hrs variable into smaller factor levels.

Currently <1, 1-2, 2-4, 4-6, 6-8, 8-10, 10-15, 15-20, 20-30, 30-40, and over 50.

I'll collapse these levels to <4, 4-10, 10-15, and over 20

```{r fct_collapse}
soc_care12$total_hrs <- 
  fct_collapse(soc_care12$total_hrs, 
               `<4` = c("<1", "1-2", "2-4"),
               `4-10` = c("4-6", "6-8", "8-10"), #10-15 stays as is
               over20 = c("20-30", "30-40", "40-50", "over50"))
```

Now summarise the data

```{r summarise_home_care_hrs}
home_care_hrs_council <-
  soc_care12 %>%
  filter(hc_client == "Yes") %>%      #keep only home care clients
  filter(age_grp != "18-<65") %>%     #keep only those over 65
  filter(council != "Highland") %>%   #remove highland due to missing data
  group_by(council, total_hrs) %>%
  summarize(N = n()) %>%
  mutate(freq = N / sum(N),
         pct = round((freq*100), 1)) 

home_care_hrs_total <-
  soc_care12 %>% 
  filter(hc_client == "Yes") %>%      #keep only home care clients
  filter(age_grp != "18-<65") %>%     #keep only those over 65
  filter(council != "Highland") %>%   #remove highland due to missing data
  group_by(total_hrs) %>% 
  summarise(N = n()) %>% 
  mutate(freq = N / sum(N),
         pct = round((freq * 100), 2))

home_care_hrs_council
home_care_hrs_total
  
```


'''and plot

```{r home_care_hrs, fig.width=16, fig.height=16}
ggplot(home_care_hrs_council, aes(total_hrs, freq)) +
  geom_bar(stat = "identity", data = home_care_hrs_total, fill = ubdc_palette[1]) +
  geom_bar(stat = "identity", fill = ubdc_palette[5], alpha = 0.6) +
  facet_geo(~ council, grid = "scotland_local_authority_grid1") +
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8)) +
  labs(
    title = "Percentage of hours of home care",
    subtitle = "by Local Authority",
    x = "",
    y = "")
  
```
























