---
title: "Social Care Survey Open Data"
subtitle: "Simple descriptive analysis"
output: 
  rmarkdown::html_document :
    theme: readable
    code_folding: hide
    toc: yes
    toc_float: yes
    number_sections: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
knitr::opts_knit$set(root.dir = 'C:/GitHub/soc_care_analysis')
```

#Introduction

Here I will use the newly released `skimr` package to describe each of the three social care datasets. Generally looking for distributions and missing data levels. 

The majority of variables are factors so we probabaly won't get the full benefit of `skimr` but am keen to try it out. 

##Load data
Load in tidied data 
```{r load_data, cache=FALSE}
load("produced_data/created_objects/soc_care10.rds")
load("produced_data/created_objects/soc_care11.rds")
load("produced_data/created_objects/soc_care12.rds")
```

##Load packages
And required packages
```{r package, message=FALSE, warning=FALSE}
library(skimr)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(readr)
library(purrr)
library(stringr)
library(tidyr)
library(magrittr)
library(forcats)
library(DT)
```


#Examine data

##Summaries using 'skimr`

1st thing is to create skim objects for each year. I've printed the 2011 versions here as an example

```{r skim_10}
skim10 <- skim(soc_care10)
skim11 <- skim(soc_care11)
skim12 <- skim(soc_care12)

datatable(skim11)
skim11
```

Here we have both the `datatable()` and console version of the skim object. The former is a tidy object that can be manipulated any way we want - the latter is a handy summary of the dataset - I particularly like the counts of complete and missing cases.

###sjmisc

Alternative 16/08/2017

```{r sjmisc}
library(sjmisc)

knitr::kable(descr(soc_care10))
```

##Councils

1st thing I want to visualise is the absolute number of records returned for each council and what percentage of the total this is. 

I'm going to stick with 2011 data for now. 

```{r skim_coucil11, fig.width=12, fig.height=8}
skim11 %>%
  filter(var == "council") %>%
  filter(stat == "count") %>%
  filter(!is.na(level)) %>%
  mutate(percentage = (value/sum(value)* 100)) %>%
  arrange(-value) %>%
  mutate(level = factor(level, level)) %>%
  select(level, value, percentage) %>%
  ggplot() +
  geom_col(aes(x = level, y = value)) + 
    geom_text(aes(label=paste0(round(percentage, 1),"%"), x= level,
                            y= value), size=4, vjust = -0.5, colour = "black") +
    theme(axis.text.x = element_text(angle = 45, size = 12, hjust = 1, vjust = 1)) +
    theme_hc() +
  labs(
    title = "Number of records returned for Social Care Survey 2011, with percentage of total",
    subtitle = "by Local Authority", 
    x = "Local Authority",
    y = "Number of records")
```


# Population comparison

How do these absolute numbers reflect the differing populations in each local authority?

##Create main table

2011 Popoulation estimates downloaded from the National Records of Scotland here:-  <https://www.nrscotland.gov.uk/statistics-and-data/statistics/statistics-by-theme/population/2011-census-reconciliation-report/list-of-tables>

Create `pop_data_combined` a `data_frame` with population sizes of each Local Authority as a total and for over 65s

```{r import_pop_data, warning=FALSE, message=FALSE}
#Note I cheated and converted data to numeric in excel to remove the commas as 1000 seperators

#import and tidy raw data
pop_data <- read_csv("raw_data/2011-cen-rec-report-pop-tab1b.csv", skip = 3) %>%
  slice(1:34) %>%
  filter(X1 %in% c("Council areas", "SCOTLAND") == FALSE) %>%
  select(-X3, -X23, -X24) %>%
  map_at(2:21, as.numeric) %>% #coerce from character to numeric
  as_data_frame()

names(pop_data) <- str_replace(names(pop_data), "X1", "council") #rename X variables
names(pop_data) <- str_replace(names(pop_data), "X2", "total")

#tidy data for manipulation
pop_data %<>%
  gather(age_grp, value, 2:21) %>%  
  arrange(council)


#create table of total population per council
pop_data_total <-
  pop_data %>%
  filter(age_grp == "total") %>%  
  select(-age_grp) %>%
  arrange(council)
names(pop_data_total) <- str_replace(names(pop_data_total), "value", "pop_total")

#create table of total over65s per council
pop_data_65plus <-
  pop_data %>%
  filter(age_grp %in% c("65 - 69", "70 - 74", "75 - 79", "80 - 84", "85 - 89", 
                        "90+") == TRUE) %>%
  group_by(council) %>%
  mutate(pop_over_65 = sum(value)) %>%    
  arrange(council) %>%
  select(council, pop_over_65) %>%
  unique(.)

#join tables together
pop_data_combined <-
  full_join(pop_data_total, pop_data_65plus)  

pop_data_combined

rm(list = c("pop_data_65plus", "pop_data_total"))
```

##Add social care

Add a column with total amount of records returned from each council i.e. the total number of indiviudals receiving some form of social care

```{r skim_11_subset}

#count number of records from each council
any_care_11 <- fct_count(soc_care11$council)

#rename tibble columns
names(any_care_11) <- 
  str_replace(names(any_care_11), "n", "any_care_total") %>%
  str_replace(., "f", "council")

#join this table to pop_data_combined
pop_data_combined %<>%
  full_join(., any_care_11)

pop_data_combined

rm(any_care_11)
```

##Add social care over 65s

Add a column with total amount of records of over 65s from each local authority

```{r subset_over_65}
#subset over 65s from 2011 data
any_care65s <-
  soc_care11 %>%
  filter(age_grp != "18-<65") 

#now count by local authority
care65s <- fct_count(any_care65s$council)
names(care65s) <- 
  str_replace(names(care65s), "n", "any_care_over_65s") %>%
  str_replace(., "f", "council")

#and add this to pop_data_combined
pop_data_combined %<>%
  full_join(., care65s)

pop_data_combined

rm(care65s)
```

##calculate ratios per 1000 population

```{r calc_per_thousand}
#calculate per 1000
pop_data_combined %<>%
  mutate(total_per_thousand = (any_care_total / pop_total) * 1000) %>%
  mutate(over_65s_per_thousand = (any_care_over_65s / pop_over_65) * 1000)

pop_data_combined
```

##Visualise over 65s

```{r vis_pop, fig.width=12, fig.height=8}

pop_data_combined %>%
  arrange(-over_65s_per_thousand) %>%
  mutate(council = factor(council, council)) %>%
  ggplot() +
  geom_col(aes(x = council, y = over_65s_per_thousand)) +
  theme(axis.text.x = element_text(angle = 45, size = 12, hjust = 1, vjust = 1)) +
  theme_hc() +
  labs(
    title = "Number of over 65s per 1000 population receiving any form of social care, 2011",
    x = "Local Authority area",
    y = "rate per 1000 population"
  )
```

##Highland

```{r check_highland}
highland <- 
  soc_care11 %>%
  filter(council == "Highland")

nrow(highland)
fct_count(highland$age_grp)
```

Highland is an outlier because `r round(2344/nrow(highland) * 100, 1)`% of its records have missing data for Age group. It is likely most of this missing data will be over 65s.


```{r skim_highland}
skim11_highland <- 
  skim(highland) %>%
  filter(stat == "missing") %>%
  filter(value != 0)

skim11_highland

```

Turns out (most likely) the same `r round(2344/nrow(highland) * 100, 1)`% of records are also missing data for `client-grp`, `gender`, and `living_arr`.

##Visualise totals

```{r vis_totals, fig.width=12, fig.height=8}
pop_data_combined %>%
  arrange(-total_per_thousand) %>%
  mutate(council = factor(council, council)) %>%
  ggplot() +
  geom_col(aes(x = council, y = total_per_thousand)) +
  theme(axis.text.x = element_text(angle = 45, size = 12, hjust = 1, vjust = 1)) +
  theme_hc() +
  labs(
    title = "Number per 1000 population receiving any form of social care, 2011",
    x = "Local Authority area",
    y = "rate per 1000 population"
  )
```

##save created objects
```{r save_obs, eval = FALSE}
save(pop_data, file = "produced_data/created_objects/pop_data.rds")
save(pop_data_combined, file = "produced_data/created_objects/pop_data_combined.rds")
save(skim10, file = "produced_data/created_objects/skim10.rds")
save(skim11, file = "produced_data/created_objects/skim11.rds")
save(skim12, file= "produced_data/created_objects/skim12.rds")
save(any_care65s, file = "produced_data/created_objects/any_care65s.rds")
```

#Variable by variable visualisation

##Councils

## Age Group

## Client Group

## Gender


## Meals

## Local Authority Provided Home Care Hours

## Private Company Provided Home Care Hours

## Voluntary Orgnaisation Provided Home Care Hours

## Personal Care Home Care Hours

## Total Home Care Hours

## Home Care Client Flad

## Living Arrangements

## Mutli Staffing

## Housing Type

## Laundry

## Shopping

## Housing Support

## IoRN Score

#Interesting comparison