---
title: "Social Care Survey Open Data"
subtitle: "Simple descriptive analysis"
output: 
  html_document:
    theme: spacelab
    toc: yes
    toc_float: yes
    number_sections: yes
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/hedgi/OneDrive - University of Glasgow/GitHub/social_care_open_data')
```

#Introduction

##Load packages

```{r packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(magrittr)
library(sf)
library(readxl)
library(RColorBrewer)
library(extrafont)
library(classInt)
theme_set(theme_minimal(base_size = 16, base_family = "Roboto"))
```


##Load data

I'm going to use the `per_thousand_table` created in `per_thousand_la_comparison.Rmd` and a shapefile downloaded from the ONS Geoportal


```{r load_data}
load("produced_data/created_objects/per_thousand_table.rds")
scotland <- st_read("raw_data/la_shapefiles/Local_Authority_Districts_December_2017_Full_Clipped_Boundaries_in_Great_Britain.shp")
la_lookup <- read_xlsx("raw_data/00534447.xlsx", sheet = "SIMD16 DZ look-up data")
```


#Data wrangle

#LA lookup data

```{r wrangle_lookup}
la_lookup %<>% 
  select(LAname) %>% 
  distinct #Keep unique names 

#Change Western Isles to match shapefile
la_lookup$LAname <- str_replace(la_lookup$LAname, "Na h-Eileanan an Iar", "Na h-Eileanan Siar") 
```

##Shape file

The shape file has every local authority in the UK - I just want Scottish Local Authorities. 

```{r wrangle_shape}
scotland %<>%
  filter(lad17nm %in% la_lookup$LAname) %>% 
  rename(council = lad17nm)
```

```{r save_shapes}
save(scotland, file = "/Users/hedgi/OneDrive - University of Glasgow/GitHub/scotland_la_shp.rds")
```

##Data

Some differences in names here also

```{r wrangle_data}
per_thousand_table$council <- str_replace(per_thousand_table$council, "Edinburgh, City of", "City of Edinburgh")
per_thousand_table$council <- str_replace(per_thousand_table$council, "Eilean Siar", "Na h-Eileanan Siar")
per_thousand_table$council <- str_replace_all(per_thousand_table$council, "&", "and")

```

#Join data to shape file

```{r join}
per_thousand_table %<>%
  full_join(., scotland, by = "council")
rm(scotland)
rm(la_lookup)
```

#Plot

#Create classes for home care

*not used*

```{r label, options}
classes <- classIntervals(per_thousand_table$homecare_per_thousand, n = 5, style = "jenks")
per_thousand_table %<>%
  mutate(homecare_class = cut(homecare_per_thousand, breaks = classes$brks, include.lowest = TRUE))
```

##Any social care

```{r plot, fig.height=10, fig.width=9}
ggplot(per_thousand_table) +
  geom_sf(aes(fill = any_care_per_thousand),
              colour = "black",
              size = 0.3) +
  scale_fill_distiller(palette = "Blues",
                      direction = 1) +
  theme(line = element_blank(),
        axis.text = element_blank(),
        panel.grid = element_line(colour = "transparent"),
        legend.position = "bottom",
        legend.text = element_text(size = 7),
        plot.caption = element_text(size = 10, colour = "#AAAAAA")) +
  labs(title = "Rate of social care for over 65s 2011/12",
       subtitle = "by Local Authority",
       fill = "Rate per thousand people\nover 65 receiving any form\nof social care",
       caption = "Contains National Statistics data © Crown copyright and database right 2018\nContains OS data © Crown copyright and database right 2018") -> any_care_map
any_care_map
```

##Home care only

```{r plot_homecare, fig.height=10, fig.width=9}
ggplot(per_thousand_table) +
  geom_sf(aes(fill = homecare_per_thousand),
              colour = "black",
              size = 0.3) +
  scale_fill_distiller(palette = "Blues",
                      direction = 1) +
  theme(line = element_blank(),
        axis.text = element_blank(),
        panel.grid = element_line(colour = "transparent"),
        legend.position = "bottom",
        legend.text = element_text(size = 7),
        plot.caption = element_text(size = 10, colour = "#AAAAAA")) +
  labs(title = "Rate of homecare for over 65s 2011/12",
       subtitle = "by Local Authority",
       fill = "Rate per thousand people\nover 65 receiving homecare",
       caption = "Contains National Statistics data © Crown copyright and database right 2018\nContains OS data © Crown copyright and database right 2018") -> homecare_map
homecare_map
```


##Telecare and alarms

```{r plot_tele, fig.height=10, fig.width=9}
ggplot(per_thousand_table) +
  geom_sf(aes(fill = telecare_per_thousand),
              colour = "black",
              size = 0.3) +
  scale_fill_distiller(palette = "Blues",
                      direction = 1) +
  theme(line = element_blank(),
        axis.text = element_blank(),
        panel.grid = element_line(colour = "transparent"),
        legend.position = "bottom",
        legend.text = element_text(size = 7),
        plot.caption = element_text(size = 10, colour = "#AAAAAA")) +
  labs(title = "Rate of telecare and community alarms\nfor over 65s 2011/12",
       subtitle = "by Local Authority",
       fill = "Rate per thousand people\nover 65 receiving any\ntelecare or community alarm\nservices",
       caption = "Contains National Statistics data © Crown copyright and database right 2018\nContains OS data © Crown copyright and database right 2018") -> telecare_map
telecare_map
```




#Save

```{r save, eval=FALSE}
ggsave("plots/anycare_map.png", any_care_map, height = 10, width = 9, dpi = 600)
ggsave("plots/homecare_map.png", homecare_map, height = 10, width = 9, dpi = 600)
ggsave("plots/telecare_map.png", telecare_map, height = 10, width = 9, dpi = 600)
```


#Session Info

```{r session_info}
devtools::session_info()
```


